A fully labelled proof system for intuitionistic modal logics

### 2021-01-15 FullyLabelled/JLC/document.tex
proof read up to end of section 3, will resume later today
final read through

@@

\begin{theorem}\label{thm:cutfree-compl}
	For any formula $\fm A$, the following are equivalent.
	%
	\begin{enumerate}
		\item\label{i} $\fm A$ is a theorem of $\IK$. 
		%
		\item\label{ii} $\fm A$ is provable in $\labIKp +\labrn{cut}$.
		\item\label{iii} $\fm A$ is provable in $\labIKp$.
		%
		\item\label{iv} $\fm A$ is valid in every birelational frame. 
	\end{enumerate}
\end{theorem}

@@

In this section we show our system at work, as most of the section
consists of derivations of axioms of $\IK$ in $\labIKp$. More precisely, we prove completeness of $\labIKp+\rn{cut}$, i.e., the implication \ref{i}$\implies$\ref{ii} of Theorem~\ref{thm:cutfree-compl}, which is stated again below:

@@

	\end{itemize}
	Next, we have to prove all axioms of intuitionistic propositional logic can be shown in $\labIKp$. We do this only for $\fm{A \AND B \IMP B}$ and leave the rest to the reader:
	\begin{smallequation*}
		\vlderivation {
			\vlin{\rlabrn{\IMP}}
			{\lb y \mbox{ fresh}}
			{\SEQ \labels{x}{A \AND B \IMP B}}
			{\vlin {\llabrn{\AND}}
				{}
				{\lseq{\futs xy}{\labels{y}{A \AND B}}{\labels{y}{B}}}
				{\vlin {\color{red}\rn{refl}}
					{}
					{\lseq{\futs xy}{\labels{y}{A}, \labels{y}{B}}{\labels{y}{B}}}
					{\vlin {\rn{id_g}}
						{}
						{\lseq{\futs xy, \futs yy}{\labels{y}{A}, \labels{y}{B}}{\labels{y}{B}}}
						{\vlhy {}}}}}
		}
	\end{smallequation*}

	Finally, we have to show how the rules of modus ponens and
	necessitation can be simulated in our system. For modus ponens, this
	is standard using the cut rule and for necessitation, we can
	transform a proof of $\fm A$ into a proof of $\fm{\BOX A}$ as
	follows.
	A proof of $\fm A$ is in fact a proof $\DD$ of the sequent
	$\SEQ{\labels{z}{A}}$ for some label $\lb z$. If $\lb x$ and $\lb y$
	are fresh labels, we can transform $\DD$ into a proof $\DD'$ of the
	sequent $\futs xy, \accs yz \SEQ \labels{z}{A}$ by adding $\futs xy,
	\accs yz$ to every line. We can now apply the $\rlabrn\BOX$-rule to
	obtain a proof of $\SEQ \labels{x}{\BOX A}$.

	This completes the proof of Theorem~\ref{thm:completeness}.
\end{proof}

Note that we could also have proved completeness via proof search and a countermodel construction argument, as rules are terminationg in $\labIKp$. 
%
This works well for $\IK$ (and a small number of its extensions) and have been used for example by~\cite{maffezioli:etal:synthese13}. 
%
Moreover, it would allow us to show completeness directly for any labelled sequents, not only specifically for formulas.

@@

In order to prove the implication \ref{iii}$\implies$\ref{iv} from
Theorem~\ref{thm:cutfree-compl} we need to show that each sequent rule
of our system $\labIKp$ is sound. To make precise what that actually
means, we have to extend the relation $\Vdash$, defined in
Section~\ref{sec:intmod} from formulas to sequents. This is the
purpose of the following definitions.

\begin{definition}\label{def:force-seq}
	Let $\M = \langle W, \rel_\M, \le_\M, V \rangle$ be a model, and let
	$\SG$ be the sequent $\lseq\B\Left\Right$. 
	%
	A \emph{$\SG$-interpretation
		in $\M$} is a mapping $\inter{\cdot}$ from the labels in $\SG$ to the set $W$ of worlds in $\M$, such that whenever $\accs xy$ in $\B$, then $\inter x\rel_\M\inter y$, and whenever $\futs xy$ in $\B$, then $\inter x\le_\M\inter y$. Now we can define 
	\begin{equation}
	\M,\inter\cdot\Vdash\SG\qquad\mbox{iff}\qquad
	\parbox{20em}{if for all $\labels{x}{A} \in \Gamma$, we have $\M, \inter x \Vdash \fm A$, then there exists $\labels{z}{B} \in \Delta$ such that $\M, \inter z \Vdash \fm B$.}
	\end{equation}
\end{definition}

\begin{definition}\label{def:valid-seq}
	A sequent $\SG$ is \emph{satisfied} in $\M = \langle W, \rel, \le, V
	\rangle$ iff for all $\SG$-interpretations $\inter\cdot$ we have
	$\M,\inter\cdot\Vdash\SG$.
	%%
	A sequent $\SG$ is \emph{valid} in a frame $\F = \langle W, \rel,
	\le \rangle$, if for all valuations $V$, the sequent $\SG$ is
	satisfied in $\langle W, R, \le, V \rangle$.
\end{definition}

@@

\begin{proof}
	We proceed by induction on the height of the derivation of $\SG$, and we show for all rules in $\labIKp$
	$$
	\vliiinf{\labrn{r}}{}{\SG}{\SGi1}{\cdots}{\SGi n}
	$$ for $n\in\{0,1,2\}$, that whenever $\SGi1,\ldots,\SGi n$ are
	valid in all birelational frame, then so is $\SG$. It follows from a case analysis on $\rn r$:
	\begin{itemize}
		\item $\rlabrn\bot$: This is trivial because $\bot$ is never forced.
		\item $\rn{id}$: This follows immediately from Proposition~\ref{prop:mon-adm}.
		\item $\llabrn\BOX$: By way of contradiction, assume that $\B, \futs xy, \accs yz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right$
		is valid in all birelational frames, but $\B, \futs xy,	\accs yz,\Left, \labels{x}{\BOX A} \SEQ \Right$ is not. This means that
		we have a model $\M$ and an interpretation $\inter\cdot$, such
		that $\M,\inter\cdot\not\Vdash\B, \futs xy, \accs yz, \Left,	\labels{x}{\BOX A} \SEQ \Right$, i.e., $\inter x\le_\M\inter y$
		and $\inter y\rel_\M\inter z$ and $\cforce \M x{\BOX A}$ but
		$\cnforce \M wB$ for all $\labels wB\in\Right$. However, by the
		definition of forcing in~\eqref{eq:kripke} we also have $\cforce
		\M zA$, and consequently $\nforce \M{\inter\cdot}{\B, \futs xy, \accs yz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ
			\Right}$. Contradiction.
		%
		\item $\rlabrn\BOX$: By way of contradiction, assume that $\B, \futs xy,
		\accs yz, \Left \SEQ \Right, \labels{z}{A}$ is valid in all
		birelational frames, but $\B, \Left \SEQ \Right, \labels{x}{\BOX
			A}$ is not, where $\lb y$ and $\lb z$ do not occur in $\B$ or
		$\Left$ or $\Right$.
		This means that we have a model $\M$ and an
		interpretation $\inter\cdot$, such that
		$\nforce\M{\inter\cdot}{\B, \Left \SEQ \Right, \labels{x}{\BOX A}}$. 
		%
		So in particular, there are worlds $\lb{y'}$ and $\lb{z'}$
		in $\M$ such that $\inter x\le_\M \lb{y'}$ and
		$\lb{y'}\rel_\M\lb{z'}$ and $\nforce\M{\lb{z'}}{\fm A}$. 
		%
		Now we let
		$\inter\cdot'$ be the extension of $\inter\cdot$ such that $\inter
		y'=\lb{y'}$ and $\inter z'=\lb{z'}$ and $\inter\cdot'=\inter\cdot$
		on all other labels. Then $\nforce\M{\inter\cdot'}{\B, \futs xy,
			\accs yz, \Left \SEQ \Right, \labels{z}{A}}$. Contradiction.

		\item $\rlabrn\IMP$: By way of contradiction, assume that $\B, \futs xy, \Left, \labels{y}{A} \SEQ \Right, \labels{y}{B}$ is valid in all birelational frames, but $\B, \Left \SEQ \Right, \labels{x}{A \IMP B}$ is not, where $\lb y$ does not occur in $\B$ or $\Left$ or $\Right$. This means that we have a model $\M$ and an interpretation $\inter\cdot$, such that $\nforce\M{\inter\cdot}{\B, \Left \SEQ \Right, \labels{x}{A \IMP B}}$. So there exists a world $\lb{y'}$ in $\M$ such that $\inter x\le_\M \lb{y'}$ and $\cforce\M{y'}{A}$ but $\nforce\M{\lb{y'}}{\fm B}$. Now we let 	$\inter\cdot'$ be the extension of $\inter\cdot$ such that $\inter y'=\lb{y'}$ and $\inter\cdot'=\inter\cdot$ on all other labels. Then $\nforce\M{\inter\cdot'}{\B, \futs xy, \Left, \labels{y}{A} \SEQ \Right, \labels{y}{B}}$. Contradiction.

@@

	\end{itemize}
	The other cases are similar (and simpler), and we leave them to the
	reader. In particular, note that the cases for the rules $\rn{refl}$,
	$\rn{trans}$, $\rn{F_1}$ and $\rn{F_2}$ are trivial, as all
	birelations frames have to obey the corresponding conditions.
\end{proof}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{theorem}
	\label{thm:cut-adm}
	The $\labrn{cut}$ rule is admissible for $\labIKp$.
\end{theorem}

@@

The next lemma looks like a special case of
Proposition~\ref{prop:mon-adm}, but it is not. First, we need to
preserve the height, and second, we cannot prove it using $\rn{cut}$ rule as we
are trying to eliminate it from derivations.

@@

\begin{proof}
	We show the details for $\llabrn{mon_a}$, the argument is the same for $\rlabrn{mon_a}$.
	By induction on the height of $\D$, we prove that for any proof of $\lseq{\B, \futs{x}{y}}{\Left, \labels{x}{a}, \labels{y}{a}}{\Right}$, there exists a proof of $\lseq{\B, \futs{x}{y}}{\Left, \labels{x}{a}}{\Right}$
	of the same (or smaller) height.
	The inductive step is straightforward by permutation of rules.
	The base cases are obtained as follows:
	\begin{smallequation*}
		\vlderivation{
			\vlin{\llabrn{mon_a}}{}{\B, \futs{x}{y}, \futs{y}{z}, \Left, \labels{x}{a} \SEQ \Right, \labels{z}{a}}{
				\vlin{\labrn{id}}{}{\B, \futs{x}{y}, \futs{y}{z}, \Left, \labels{x}{a}, \labels{y}{a} \SEQ \Right, \labels{z}{a}}{
					\vlhy{}
				}
			}
		}
		\reducesto
		\vlderivation{
			\vlin{\color{red}\labrn{trans}}{}{\B, \futs{x}{y}, \futs{y}{z}, \Left, \labels{x}{a} \SEQ \Right, \labels{z}{a}}{\vlin{\labrn{id}}{}{\B, \futs{x}{y}, \futs{y}{z}, \futs{x}{z}, \Left, \labels{x}{a} \SEQ \Right, \labels{z}{a}}{
					\vlhy{}
				}}
			}
		\end{smallequation*}

		\begin{smallequation*}
			\vlderivation{
				\vlin{\llabrn{mon_a}}{}{\B, \futs{x}{y}, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a}}{
					\vlin{\labrn{id}}{}{\B, \futs{x}{y}, \labels{x}{a}, \labels{y}{a} \SEQ \Right, \labels{x}{a}}{
						\vlhy{}
					}
				}
			}
			\reducesto
			\vlderivation{
				\vlin{\color{red}\labrn{refl}}{}{\B, \futs{x}{y}, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a}}{\vlin{\labrn{id}}{}{\B, \futs{x}{y}, \futs{x}{x}, \labels{x}{a} \SEQ \Right, \labels{x}{a}}{
						\vlhy{}
					}}}
				\end{smallequation*}

				\begin{smallequation*}
					\vlderivation{
						\vlin{\llabrn{mon_a}}{}{\B, \futs{x}{y}, \Left, \labels{x}{a} \SEQ \Right, \labels{y}{a}}{
							\vlin{\labrn{id}}{}{\B, \futs{x}{y}, \labels{x}{a}, \labels{y}{a} \SEQ \Right, \labels{y}{a}}{
								\vlhy{}
							}
						}
					}
					\reducesto
					\vlinf{\labrn{id}}{}{\B, \futs{x}{y}, \Left, \labels{x}{a} \SEQ \Right, \labels{y}{a}}{}
					\qedhere
				\end{smallequation*}

@@

			The next lemma shows that all the rules in our system are invertible, as already mentioned in the introduction.

			\begin{lemma}
				\label{lem:inv}
				All single-premise rules of $\labIKp$ are height-preserving
				invertible. Furthermore, the rules $\llabrn{\OR}$ and
				$\rlabrn{\AND}$ are height-preserving invertible on both premises, and the
				rule $\llabrn{\IMP}$ is height-preserving invertible on the right premise.
				%
			\end{lemma}

			\begin{proof}
				For each rule $\rn{r}$, we need to show that if there exists a proof $\DD$ of the conclusion, there exists a proof $\DD^{\rn r_i}$ of the $i$-th premise, of the same (or smaller) height.
				For $\rlabrn\AND$, $\llabrn\AND$, $\rlabrn\OR$, $\llabrn\OR$, and the right premise of $\llabrn\IMP$, we use a standard induction on the height of $\DD$.
				%
				For $\rlabrn\IMP$, $\rlabrn\BOX$, $\llabrn\DIA$ as well, but
				we need to make sure that the obtained derivation uses a fresh label by using substitution inside $\DD^{\rn r}$ when necessary.
				%
				The other rules can be shown invertible using Lemma~\ref{lem:weak-adm}. 
			\end{proof}

@@

				\begin{description}
					\item[Base cases:]\label{base-cases}
					When the last rule in $\DD_1$ is an axiom, 
					%
					we can produce directly a cut-free derivation of the conclusion.
					%
					In the first case, we appeal to Lemma~\ref{lem:adm-mon-at}, to use the atomic monotonicity rule freely.
					\begin{itemize}
						\item 
						\begin{smallequation*}
							\vlderiibase{\labrn{cut}}{}{\B, \futs xy, \Left, \labels{x}{a} \SEQ \Right}{
								\vlin{\labrn{id}}{}{\B, \futs xy, \Left, \labels{x}{a} \SEQ \Right, \labels{y}{a}}{
									\vlhy{}	
								}
							}{
							\vlhtr{\DD_2}{\B, \futs xy, \Left, \labels{x}{a}, \labels{y}{a} \SEQ \Right}		
						}
						\reducesto
						\vlderibase{\llabrn{mon_a}}{}{\B, \futs xy, \Left, \labels{x}{a} \SEQ \Right}{
							\vlhtr{\DD_2^{\rn w}}{\B, \futs xy, \Left, \labels{x}{a}, \labels{y}{a} \SEQ \Right}
						}
					\end{smallequation*}

@@

	\begin{smallequation*}
		\reducesto\qquad\vlderiibase{\llabrn\IMP}{}{\B, \futs xy, \Left, \labels{x}{A \IMP B} \SEQ \Right}{
			\vlhtr{\DD_1'}{\B, \futs xy, \Left, \labels{x}{A \IMP B} \SEQ \Right, \labels{y}{A}}
		}{
		\vlhtr{\DD_2'}{\B, \futs xy, \Left, \labels{y}{B} \SEQ \Right}
	}
\end{smallequation*}

\begin{small}
	{with}$\quad\DD_1' = \quad
	\vlderiibase{\labrn{cut}}{}{\B, \futs xy, \Left, \labels{x}{A \IMP B} \SEQ \Right, \labels{y}{A}}{
		\vlhtr{\DD_1}{\B, \futs xy, \Left, \labels{x}{A \IMP B} \SEQ \Right, \labels{z}{C}, \labels{y}{A}}
	}{
	\vlhtr{\DD_3^{\rn w}}{\B, \futs xy, \Left, \labels{x}{A \IMP B}, \labels{z}{C} \SEQ \Right, \labels{y}{A}}
}	$

{and}$\quad\DD_2'= \quad
\vlderiibase{\labrn{cut}}{}{\B, \futs xy, \Left, \labels{y}{B} \SEQ \Right}{
	\vlhtr{\DD_2}{\B, \futs xy, \Left, \labels{y}{B} \SEQ \Right, \labels{z}{C}}
}{
\vlhtr{\DD_3^{\invr{\llabrn\IMP}}}{\B, \futs xy, \Left, \labels{y}{B}, \labels{z}{C} \SEQ \Right}
}$
\end{small}

>>>>

\newcommand{\set}[1]{\{#1\}}

@@

\begin{theorem}\label{thm:cutfree-compl}
	For any formula $\fm A$, the following are equivalent.
	%
	\begin{enumerate}
		\item\label{i} $\fm A$ is a theorem of $\IK$. 
		%
		\item\label{ii} $\fm A$ is provable in $\labIKp +\labrn{cut}$.
		%
		\item\label{iii} $\fm A$ is provable in $\labIKp$.
		%
		\item\label{iv} $\fm A$ is valid in every birelational frame. 
	\end{enumerate}
\end{theorem}

@@

In this section we show our system at work, as most of the section
consists of derivations of axioms of $\IK$ in $\labIKp$. More precisely, we prove completeness of $\labIKp+\rn{cut}$, i.e., the implication \ref{i}$\implies$\ref{ii} of Theorem~\ref{thm:cutfree-compl}, which is stated again below.

@@

	\end{itemize}
	Next, we have to prove that all axioms of intuitionistic propositional logic can be shown in $\labIKp$. We do this only for $\fm{A \AND B \IMP B}$ and leave the rest to the reader.
	\begin{smallequation*}
		\vlderivation {
			\vlin{\rlabrn{\IMP}}
			{\lb y \mbox{ fresh}}
			{\SEQ \labels{x}{A \AND B \IMP B}}
			{\vlin {\llabrn{\AND}}
				{}
				{\lseq{\futs xy}{\labels{y}{A \AND B}}{\labels{y}{B}}}
				{\vlin {\color{red}\rn{refl}}
					{}
					{\lseq{\futs xy}{\labels{y}{A}, \labels{y}{B}}{\labels{y}{B}}}
					{\vlin {\rn{id_g}}
						{}
						{\lseq{\futs xy, \futs yy}{\labels{y}{A}, \labels{y}{B}}{\labels{y}{B}}}
						{\vlhy {}}}}}
		}
	\end{smallequation*}

	Finally, we have to show how the rules of modus ponens and
	necessitation can be simulated in our system. For modus ponens, it
	is standardly done using the cut rule, and for necessitation, we can
	transform a proof of $\fm A$ into a proof of $\fm{\BOX A}$ as
	follows.
	A proof of $\fm A$ is in fact a proof $\DD$ of the sequent
	$\SEQ{\labels{z}{A}}$ for some label $\lb z$. If $\lb x$ and $\lb y$
	are fresh labels, we can transform $\DD$ into a proof $\DD'$ of the
	sequent $\futs xy, \accs yz \SEQ \labels{z}{A}$ by adding $\futs xy,
	\accs yz$ to every line. We can now apply the $\rlabrn\BOX$-rule to
	obtain a proof of $\SEQ \labels{x}{\BOX A}$.
%	
%	This completes the proof of Theorem~\ref{thm:completeness}.
\end{proof}

Note that we could also have proved completeness via proof search and a countermodel construction argument, as rules are terminationg in $\labIKp$. 
%
This works well for $\IK$ (and a small number of its extensions) and has been used for example by~\cite{maffezioli:etal:synthese13}. 
%
Moreover, it would allow us to show completeness directly for any labelled sequents, not only specifically for formulas.

@@

In order to prove the implication \ref{iii}$\implies$\ref{iv} from
Theorem~\ref{thm:cutfree-compl} we need to show that each sequent rule
of our system $\labIKp$ is sound. To make precise what that actually
means, we have to extend the relation $\Vdash$, defined in
Section~\ref{sec:intmod}, from formulas to sequents. This is the
purpose of the following definitions.

\begin{definition}\label{def:force-seq}
	Let $\M = \langle W, \rel_\M, \le_\M, V \rangle$ be a model, and let
	$\SG$ be the sequent $\lseq\B\Left\Right$. 
	%
	A \emph{$\SG$-interpretation
		in $\M$} is a mapping $\inter{\cdot}$ from the labels in $\SG$ to the set $W$ of worlds in $\M$, such that whenever $\accs xy$ in $\B$, then $\inter x\rel_\M\inter y$, and whenever $\futs xy$ in $\B$, then $\inter x\le_\M\inter y$. Now we can define 
	\begin{equation}
	\M,\inter\cdot\Vdash\SG\qquad\mbox{iff}\qquad
	\parbox{20em}{if for all $\labels{x}{A} \in \Gamma$, we have $\M, \inter x \Vdash \fm A$, then there exists $\labels{z}{B} \in \Delta$, such that $\M, \inter z \Vdash \fm B$.}
	\end{equation}
\end{definition}

\begin{definition}\label{def:valid-seq}
	A sequent $\SG$ is \emph{satisfied} in $\M = \langle W, \rel, \le, V
	\rangle$ if for all $\SG$-interpretations $\inter\cdot$ we have
	$\M,\inter\cdot\Vdash\SG$.
	%%
	A sequent $\SG$ is \emph{valid} in a frame $\F = \langle W, \rel,
	\le \rangle$ if, for all valuations $V$, the sequent $\SG$ is
	satisfied in $\langle W, R, \le, V \rangle$.
\end{definition}

@@

\begin{proof}
	We proceed by induction on the height of the derivation of $\SG$, and we show for all rules in $\labIKp$
	$$
	\vliiinf{\labrn{r}}{}{\SG}{\SGi1}{\cdots}{\SGi n}
	$$ for $n\in\{0,1,2\}$, that whenever $\SGi1,\ldots,\SGi n$ are
	valid in all birelational frame, then so is $\SG$. It follows from a case analysis on $\rn r$:
	\begin{itemize}
		\item $\rlabrn\bot$: This is trivial because $\bot$ is never forced.
		\item $\rn{id}$: This follows immediately from Proposition~\ref{prop:mon-adm}.
		\item $\llabrn\BOX$: By way of contradiction, assume that $\B, \futs xy, \accs yz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right$
		is valid in all birelational frames, but $\B, \futs xy,	\accs yz,\Left, \labels{x}{\BOX A} \SEQ \Right$ is not. This means that
		we have a model $\M$ and an interpretation $\inter\cdot$, such
		that $\M,\inter\cdot\not\Vdash\B, \futs xy, \accs yz, \Left,	\labels{x}{\BOX A} \SEQ \Right$, i.e., $\inter x\le_\M\inter y$
		and $\inter y\rel_\M\inter z$ and $\cforce \M x{\BOX A}$ but
		$\cnforce \M wB$ for all $\labels wB\in\Right$. However, by the
		definition of forcing in~\eqref{eq:kripke} we also have $\cforce
		\M zA$, and consequently $\nforce \M{\inter\cdot}{\B, \futs xy, \accs yz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ
			\Right}$. Contradiction.
		%
		\item $\rlabrn\BOX$: By way of contradiction, assume that $\B, \futs xy,
		\accs yz, \Left \SEQ \Right, \labels{z}{A}$ is valid in all
		birelational frames, but $\B, \Left \SEQ \Right, \labels{x}{\BOX
			A}$ is not, where $\lb y$ and $\lb z$ do not occur in $\B$ or
		$\Left$ or $\Right$.
		This means that we have a model $\M$ and an
		interpretation $\inter\cdot$, such that
		$\nforce\M{\inter\cdot}{\B, \Left \SEQ \Right, \labels{x}{\BOX A}}$. 
		%
		In particular, there are worlds $\lb{y'}$ and $\lb{z'}$
		in $\M$ such that $\inter x\le_\M \lb{y'}$ and
		$\lb{y'}\rel_\M\lb{z'}$ and $\nforce\M{\lb{z'}}{\fm A}$. 
		%
		Now we define
		the extension $\inter\cdot'$ of $\inter\cdot$ as $\inter
		y'=\lb{y'}$ and $\inter z'=\lb{z'}$ and $\inter\cdot'=\inter\cdot$
		on all other labels. Then $\nforce\M{\inter\cdot'}{\B, \futs xy,
			\accs yz, \Left \SEQ \Right, \labels{z}{A}}$. Contradiction.

		\item $\rlabrn\IMP$: By way of contradiction, assume that $\B, \futs xy, \Left, \labels{y}{A} \SEQ \Right, \labels{y}{B}$ is valid in all birelational frames, but $\B, \Left \SEQ \Right, \labels{x}{A \IMP B}$ is not, where $\lb y$ does not occur in $\B$ or $\Left$ or $\Right$. This means that we have a model $\M$ and an interpretation $\inter\cdot$, such that $\nforce\M{\inter\cdot}{\B, \Left \SEQ \Right, \labels{x}{A \IMP B}}$. So there exists a world $\lb{y'}$ in $\M$ such that $\inter x\le_\M \lb{y'}$ and $\cforce\M{y'}{A}$ but $\nforce\M{\lb{y'}}{\fm B}$. Now let 	$\inter\cdot'$ be the extension of $\inter\cdot$ such that $\inter y'=\lb{y'}$ and $\inter\cdot'=\inter\cdot$ on all other labels. Then $\nforce\M{\inter\cdot'}{\B, \futs xy, \Left, \labels{y}{A} \SEQ \Right, \labels{y}{B}}$. Contradiction.

@@

	\end{itemize}
	The other cases are similar (and simpler), and we leave them to the
	reader. In particular, note that the cases for the rules $\rn{refl}$,
	$\rn{trans}$, $\rn{F_1}$ and $\rn{F_2}$ are trivial, as all
	birelational frames have to obey the corresponding conditions.
\end{proof}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{theorem}
	\label{thm:cut-adm}
	All the occurences of the $\labrn{cut}$ rule in a proof $\DD$ in $\labIKp$ can be eliminated.
\end{theorem}

@@

The next lemma looks like a special case of
Proposition~\ref{prop:mon-adm}, but it is not. First, we need to
preserve the height, and second, we cannot prove it using the $\rn{cut}$ rule as we
are trying to eliminate it from derivations.

@@

\begin{proof}
	We show the details for $\llabrn{mon_a}$, the argument is the same for $\rlabrn{mon_a}$.
	By induction on the height of $\D$, we prove that for any proof of $\lseq{\B, \futs{x}{y}}{\Left, \labels{x}{a}, \labels{y}{a}}{\Right}$, there exists a proof of $\lseq{\B, \futs{x}{y}}{\Left, \labels{x}{a}}{\Right}$
	of the same (or smaller) height.
	The inductive step is straightforward by permutation of rules.
	The base cases are obtained as follows:
	\begin{smallequation*}
		\vlderibase{\llabrn{mon_a}}{}{\B, \futs{x}{y}, \futs{y}{z}, \Left, \labels{x}{a} \SEQ \Right, \labels{z}{a}}{
				\vlin{\labrn{id}}{}{\B, \futs{x}{y}, \futs{y}{z}, \Left, \labels{x}{a}, \labels{y}{a} \SEQ \Right, \labels{z}{a}}{
					\vlhy{}
				}
			}
		\reducesto
		\vlderibase{\color{red}\labrn{trans}}{}{\B, \futs{x}{y}, \futs{y}{z}, \Left, \labels{x}{a} \SEQ \Right, \labels{z}{a}}{\vlin{\labrn{id}}{}{\B, \futs{x}{y}, \futs{y}{z}, \futs{x}{z}, \Left, \labels{x}{a} \SEQ \Right, \labels{z}{a}}{
					\vlhy{}
				}
			}
		\end{smallequation*}

		\begin{smallequation*} 
 			\vlderibase{\llabrn{mon_a}}{}{\B, \futs{x}{y}, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a}}{
					\vlin{\labrn{id}}{}{\B, \futs{x}{y}, \labels{x}{a}, \labels{y}{a} \SEQ \Right, \labels{x}{a}}{
						\vlhy{}
					}
				}
			\reducesto
			\vlderibase{\color{red}\labrn{refl}}{}{\B, \futs{x}{y}, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a}}{\vlin{\labrn{id}}{}{\B, \futs{x}{y}, \futs{x}{x}, \labels{x}{a} \SEQ \Right, \labels{x}{a}}{
						\vlhy{}
					}}
				\end{smallequation*}

				\begin{smallequation*}
					\vlderibase{\llabrn{mon_a}}{}{\B, \futs{x}{y}, \Left, \labels{x}{a} \SEQ \Right, \labels{y}{a}}{
							\vlin{\labrn{id}}{}{\B, \futs{x}{y}, \labels{x}{a}, \labels{y}{a} \SEQ \Right, \labels{y}{a}}{
								\vlhy{}
							}
						}
					\reducesto
					\vlinf{\labrn{id}}{}{\B, \futs{x}{y}, \Left, \labels{x}{a} \SEQ \Right, \labels{y}{a}}{}
					\qedhere
				\end{smallequation*}

@@

			The next lemma shows that the rules in our system are invertible, as already mentioned in the introduction.

			\begin{lemma}
				\label{lem:inv}
				All single-premise rules of $\labIKp$ are height-preserving
				invertible. Furthermore, the rules $\llabrn{\OR}$ and
				$\rlabrn{\AND}$ are height-preserving invertible on both premises, and the
				rule $\llabrn{\IMP}$ is height-preserving invertible on the right premise only.
				%
			\end{lemma}

			\begin{proof}
				For each rule $\rn{r}$, we need to show that if there exists a proof $\DD$ of the conclusion, there exists a proof $\DD^{\rn r_i}$ of the $i$-th premise, of the same (or smaller) height.
				For $\rlabrn\AND$, $\llabrn\AND$, $\rlabrn\OR$, $\llabrn\OR$, and the right premise of $\llabrn\IMP$, we use a standard induction on the height of $\DD$.
				%
				For $\rlabrn\IMP$, $\rlabrn\BOX$, $\llabrn\DIA$ as well, but
				we need to make sure that the obtained derivation uses a fresh label by using substitution inside $\DD^{\rn r_i}$ when necessary.
				%
				The other rules can be shown invertible using Lemma~\ref{lem:weak-adm}. 
			\end{proof}

@@

				\begin{description}
					\item[Base cases:]\label{base-cases}
					When the last rule in $\DD_1$ is an axiom, 
					%
					we can produce directly a cut-free derivation of the conclusion.
					%
					In the first case, we appeal to Lemma~\ref{lem:adm-mon-at}, to use the atomic monotonicity rule freely and to Lemma~\ref{lem:weak-adm} to obtain $\DD_2^{\rn w}$ through weakening admissibility.
					\begin{itemize}
						\item 
						\begin{smallequation*}
							\vlderiibase{\labrn{cut}}{}{\B, \futs xy, \Left, \labels{x}{a} \SEQ \Right}{
								\vlin{\labrn{id}}{}{\B, \futs xy, \Left, \labels{x}{a} \SEQ \Right, \labels{y}{a}}{
									\vlhy{}	
								}
							}{
							\vlhtr{\DD_2}{\B, \futs xy, \Left, \labels{x}{a}, \labels{y}{a} \SEQ \Right}		
						}
						\reducesto
						\vlderibase{\llabrn{mon_a}}{}{\B, \futs xy, \Left, \labels{x}{a} \SEQ \Right}{
							\vlhtr{\DD_2^{\rn w}}{\B, \futs xy, \Left, \labels{x}{a}, \labels{y}{a} \SEQ \Right}
						}
					\end{smallequation*}

@@

	\begin{smallequation*}
		\reducesto\vlderiibase{\llabrn\IMP}{}{\B, \futs xy, \Left, \labels{x}{A \IMP B} \SEQ \Right}{
			\vlhtr{\DD_1'}{\B, \futs xy, \Left, \labels{x}{A \IMP B} \SEQ \Right, \labels{y}{A}}
		}{
%		\vlhtr{\DD_2'}{\B, \futs xy, \Left, \labels{y}{B} \SEQ \Right}
			\vliin{\labrn{cut}}{}{\B, \futs xy, \Left, \labels{y}{B} \SEQ \Right}{
				\vlhtr{\DD_2}{\B, \futs xy, \Left, \labels{y}{B} \SEQ \Right, \labels{z}{C}}
			}{
				\vlhtr{\DD_3^{\invr{\llabrn\IMP}}}{\B, \futs xy, \Left, \labels{y}{B}, \labels{z}{C} \SEQ \Right}
			}
	}
\end{smallequation*}

with
\begin{small}
	$\quad\DD_1' = \quad
	\vlderiibase{\labrn{cut}}{}{\B, \futs xy, \Left, \labels{x}{A \IMP B} \SEQ \Right, \labels{y}{A}}{
		\vlhtr{\DD_1}{\B, \futs xy, \Left, \labels{x}{A \IMP B} \SEQ \Right, \labels{z}{C}, \labels{y}{A}}
	}{
	\vlhtr{\DD_3^{\rn w}}{\B, \futs xy, \Left, \labels{x}{A \IMP B}, \labels{z}{C} \SEQ \Right, \labels{y}{A}}
	}	
	$
\end{small}

%and
%\begin{small}
%$\quad\DD_2'= \quad
%\vlderiibase{\labrn{cut}}{}{\B, \futs xy, \Left, \labels{y}{B} \SEQ \Right}{
%	\vlhtr{\DD_2}{\B, \futs xy, \Left, \labels{y}{B} \SEQ \Right, \labels{z}{C}}
%}{
%\vlhtr{\DD_3^{\invr{\llabrn\IMP}}}{\B, \futs xy, \Left, \labels{y}{B}, \labels{z}{C} \SEQ \Right}
%}$
%\end{small}

 where $\DD_3^{\rn w}$ is obtained using Lemma~\ref{lem:weak-adm} and $\DD_3^{\llabrn\IMP}$ is obtained using Lemma~\ref{lem:inv}. We use the same naming scheme in the following cases.

\item $\rlabrn\IMP$:
\begin{smallequation*}
	\vlderiibase{\labrn{cut}}{}{\B, \Left \SEQ \Right, \labels{x}{A \IMP B}}{
		\vlin{\rlabrn\IMP}{\text{\footnotesize $\lb{x'}$ fresh}}{\B, \Left \SEQ \Right, \labels{x}{A \IMP B}, \labels{z}{C}}{
			\vlhtr{\DD_1}{\B, \futs{x}{x'}, \Left, \labels{x'}{A} \SEQ \Right, \labels{x'}{B}, \labels{z}{C}}
		}
	}{
	\vlhtr{\DD_2}{\B, \Left, \labels{z}{C} \SEQ \Right, \labels{x}{A \IMP B}}
}
\end{smallequation*}
